### 2023-09-01 proofs/main.tex
Finish syntax update

\subsection{Algorithmic Renaming}

In several places of our algorithm, 
we turn a declarative type into the algorithmic one
via replacing certain type variables with fresh algorithmic variables.
This is done via algorithmic renaming

\begin{definition}
  Suppose that $[[nas]]$ is a list of negative type variables,

@@

  $[[dom(Θ)]]$ denotes the set of variables appearing in $[[Θ]]$:
  $[[dom(Θ)]] = \{ [[α̂±]] \mid [[α̂±[Γ] ]] \in [[Θ]] \}$.
\end{definition}

@@

\begin{definition}[Signature of Algorithmic Substitution]
  $[[Θ ⊢ uσ : Θ]]$ means that
  \begin{enumerate}
    \item for any $[[α̂± ∊ Ξ]]$,
      there exists $[[Γ]]$ such that $[[ α̂±[Γ] ∊ Θ ]]$
      and $[[ Γ ⊢ [uσ]α̂± ]]$; 
    \item for any $[[ α̂± ∉ Ξ]]$, $[[ [uσ]α̂± ]] =  [[ α̂± ]]$.
  \end{enumerate}
\end{definition}

@@

\begin{definition}[Unification Constraint]
  \hfill
  \begin{itemize}
    \item Unification entry (denoted as $[[ucE]]$) is an expression of shape 
      $[[pua :≈ iP]]$ or $[[nua :≈ iN]]$;
    \item unification constraint (denoted as $[[UC]]$) is a set of 
      unification constraint entries.
  \end{itemize}

\end{definition}

@@

@@

Two matching entries formed in the same context $[[Γ]]$ 
can be merged in the following way:
\begin{definition}[Merge of Matching Constraint Entries]
   \hfill \\
\ottdefnSCME\\
\end{definition}
% Notice that in case of equivalence, the assigned types
% must be equal (i.e. alpha-equivalent) to be merged. This is because
% the unification algorithm assumes that every type is normalized,
% and hence, equivalence is alpha-equivalence 
% (\cref{corollary:nf-complete-wrt-subt-equiv,corollary:nf-sound-wrt-subt-equiv}).

\begin{definition}[Constraint Context]
  Constraint context (denoted as $[[Θ]]$) is a set of entries of shape $[[ pua[Γ] ]]$ 
  and $[[ nua[Γ] ]]$, specifying that the unification variable ($[[pua]]$ or $[[nua]]$)
  must be instantiated in a type well-formed in context $[[Γ]]$.
  We assume that for each $[[α̂±]]$, at most one entry of shape $[[ α̂±[Γ] ]]$ is in $[[Θ]]$.
  If $[[ α̂±[Γ] ]] \in [[Θ]]$, we denote $[[Γ]]$ as $[[Θ(α̂±)]]$.
\end{definition}

\begin{definition}[Well-formed Constraint]
  We say that a constraint is well-formed in a
  constraint context $[[Θ]]$ if all its entries are well-formed in
  the corresponding elements of $[[Θ]]$.
  More formally, 
  $[[Θ ⊢ UC]]$ iff for every $[[ucE]] \in $ $[[UC]]$,
  such that $[[ucE]]$ restricts $[[α̂±]]$,
  there exists $[[ α̂±[Γ] ]] \in [[Θ]]$
  and $[[Γ ⊢ ucE]]$.
\end{definition}

\begin{definition}[Merge of Subtyping Constraints]
  Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.
  Then $[[Θ ⊢ SC1 & SC2 = SC]]$
  defines a set such that $[[ucE]] \in [[SC]]$ iff either
  \begin{itemize}
    \item $[[ucE]] \in [[SC1]]$ and there is no matching $[[ucE']] \in [[SC2]]$; or
    \item $[[ucE]] \in [[SC2]]$ and there is no matching $[[ucE']] \in [[SC1]]$; or
    \item $[[Θ(α̂±) ⊢ ucE1 & ucE2 = ucE]]$ for some $[[ucE1]] \in [[SC1]]$ and $[[ucE2]] \in [[SC2]]$
      such that $[[ucE1]]$ matches with $[[ucE2]]$ restricting variable
      $[[α̂±]]$. 
  \end{itemize}
\end{definition}

>>>>

\subsection{Variable Algorithmization}

In several places of our algorithm, 
we turn a declarative type into the algorithmic one
via replacing certain type variables with fresh algorithmic variables.

\begin{definition}[Variable Algorithmization]
  Suppose that $[[nas]]$ is a list of negative type variables
  and $[[nuas]]$ is a list of negative algorithmic variables of the same length. 
  Then $[[ nuas/nas ]]$ is a substitution turning each $[[αi⁻]]$ from $[[nas]]$ into 
  $[[αî⁻]]$ from $[[nuas]]$.
  Positive algorithmization $[[ puas/pas ]]$
  is defined symmetrically.
\end{definition}

@@

  $[[dom(Θ)]]$ denotes the set of variables appearing in $[[Θ]]$:
  $[[dom(Θ)]] = \{ [[α̂±]] \mid [[α̂±[Γ] ]] \in [[Θ]] \}$.

  If $[[ α̂±[Γ] ]] \in [[Θ]]$, we denote $[[Γ]]$ as $[[Θ(α̂±)]]$.
\end{definition}

@@

\begin{definition}[Signature of Algorithmic Substitution]
  \hfill
  \begin{itemize}
    \item $[[Θ ⊢ uσ : Ξ]]$ means that
      \begin{enumerate}
        \item for any $[[α̂± ∊ Ξ]]$,
          there exists $[[Γ]]$ such that $[[ α̂±[Γ] ∊ Θ ]]$
          and $[[ Γ ⊢ [uσ]α̂± ]]$; 
        \item for any $[[ α̂± ∉ Ξ]]$, $[[ [uσ]α̂± ]] =  [[ α̂± ]]$.
      \end{enumerate}
    \item $[[Γ ⊢ uσ : Ξ]]$ means that
      \begin{enumerate}
        \item for any $[[α̂± ∊ Ξ]]$, $[[ Γ ⊢ [uσ]α̂± ]]$; 
        \item for any $[[ α̂± ∉ Ξ]]$, $[[ [uσ]α̂± ]] =  [[ α̂± ]]$.
      \end{enumerate}
  \end{itemize}
\end{definition}

@@

\begin{definition}[Unification Constraint]
  \hfill
  \begin{itemize}
    \item Unification entry (denoted as $[[ucE]]$) is an expression of shape 
      $[[pua :≈ iP]]$ or $[[nua :≈ iN]]$;
    \item unification constraint (denoted as $[[UC]]$) is a set of 
      unification constraint entries.
  \end{itemize}
\end{definition}

However, in subtyping we need to consider more general
kind of constraints. Specifically,
subtyping constraint entries can restrict a variable
not only to be equivalent to a certain type, but
also to be a supertype of a positive type.

@@

\begin{definition}[Well-formed Constraint]
  We say that a constraint is well-formed in a
  subtyping context $[[Θ]]$ if all its entries are well-formed in
  the corresponding elements of $[[Θ]]$.
  More formally, 
  $[[Θ ⊢ UC]]$ iff for every $[[ucE]] \in $ $[[UC]]$,
  such that $[[ucE]]$ restricts $[[α̂±]]$,
  there exists $[[ α̂±[Γ] ]] \in [[Θ]]$
  and $[[Γ ⊢ ucE]]$.
\end{definition}

Next, define the least upper bound 
for two subtyping constraints.
First, we define the least upper bound for entries,
and then extend it to the set of entries.

@@

Two matching entries formed in the same context $[[Γ]]$ 
can be merged in the following way:
\begin{definition}[Merge of Matching Constraint Entries]
  \label{definition:merge-matching-entries}
   \hfill \\
  \ottdefnSCME\\
\end{definition}

Unification constraint entries are a special case of subtyping constraint
entries. They are merged using the same algorithm 
(\cref{definition:merge-matching-entries}).
Notice that the merge of two matching unification constraint entries
is a unification constraint entry.
\begin{lemma}[Merge of Matching Unification Constraint Entries is well-defined]
  Suppose that $[[Γ ⊢ ucE1]]$ and $[[Γ ⊢ ucE2]]$
  are unification constraint entries. 
  Then the merge of $[[ucE1]]$ and $[[ucE2]]$ 
  $[[Γ ⊢ ucE1 & ucE2 = ucE]]$
  according to \cref{definition:merge-matching-entries},
  is a unification constraint entry.
\end{lemma}
\begin{proof}
  Since $[[ucE1]]$ and $[[ucE2]]$ are matching unification constraint entries,
  they have the shape $([[pua :≈ iP1]], [[pua :≈ iP2]])$ or
  $([[nua :≈ iN1]], [[nua :≈ iN2]])$.
  Then the merge of $[[ucE1]]$ and $[[ucE2]]$ 
  can only be defined by \ruleref{\ottdruleSCMEPEqEqLabel} or
  \ruleref{\ottdruleSCMENEqEqLabel}.
  In both cases the result, if it exists, 
  is a unification constraint entry:
  in the first case, the result has shape $[[pua :≈ iP1]]$,
  in the second case, the result has shape $[[nua :≈ iN1]]$.
\end{proof}

% Notice that in case of equivalence, the assigned types
% must be equal (i.e. alpha-equivalent) to be merged. This is because
% the unification algorithm assumes that every type is normalized,
% and hence, equivalence is alpha-equivalence 
% (\cref{corollary:nf-complete-wrt-subt-equiv,corollary:nf-sound-wrt-subt-equiv}).

\begin{definition}[Merge of Subtyping Constraints]
  \label{definition:merge-subtyping-constraints}
  Suppose that $[[Θ ⊢ SC1]]$ and $[[Θ ⊢ SC2]]$.
  Then $[[Θ ⊢ SC1 & SC2 = SC]]$
  defines a set such that $[[ucE]] \in [[SC]]$ iff either
  \begin{itemize}
    \item $[[ucE]] \in [[SC1]]$ and there is no matching $[[ucE']] \in [[SC2]]$; or
    \item $[[ucE]] \in [[SC2]]$ and there is no matching $[[ucE']] \in [[SC1]]$; or
    \item $[[Θ(α̂±) ⊢ ucE1 & ucE2 = ucE]]$ for some $[[ucE1]] \in [[SC1]]$ and $[[ucE2]] \in [[SC2]]$
      such that $[[ucE1]]$ matches with $[[ucE2]]$ restricting variable
      $[[α̂±]]$. 
  \end{itemize}
\end{definition}

Unification constraints can be considered 
as a special case of subtyping constraints,
and the merge of unification constraints
is defined as the merge of subtyping constraints.
Then it is easy to see that the merge of two 
unification constraints is a unification constraint.

\begin{lemma}[Merge of Unification Constraints is well-defined]
  Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$
  are unification constraints. 
  Then the merge of $[[UC1]]$ and $[[UC2]]$ 
  $[[Θ ⊢ lift UC1 & lift UC2 = lift UC]]$
  according to \cref{definition:merge-subtyping-constraints},
  is a unification constraint.
\end{lemma}
\begin{proof}
  $[[UC]]$ consists of unmatched entries of $[[UC1]]$ and $[[UC2]]$,
  which are \emph{unification} constraint entries by assumption,
  and merge of matching entries, which also are  
  \emph{unification} constraint entries by \cref{lemma:merge-matching-entries-is-well-defined}.
\end{proof}

### 2023-09-01 proofs/tex/equiv-lemmas.tex
WIP: update typing lemmas
Finish syntax update

\begin{corollary}[Normalization preserves the signature of the algorithmic substitution]
  \label{corollary:norm-subst-sig-algo}
  \hfill \\
  $[[Θ ⊢ uσ]] \iff [[Θ ⊢ nf(uσ)]]$ and
  $[[Γ ⊢ uσ : Ξ]] \iff [[Γ ⊢ nf(uσ) : Ξ]]$.
\end{corollary}
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

@@

\begin{lemma}[Equivalent substitution act equivalently]
  \label{lemma:equiv-subst-on-same-term}
  Suppose that $[[Γ' ⊢ σ1 : Γ]]$ and $[[Γ' ⊢ σ2 : Γ]]$
  are substitutions equivalent on their domain, that is $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$.
  Then
  \begin{itemize}
    \item[$+$] for any $[[Γ ⊢ iP]]$, $[[Γ' ⊢ [σ1]iP ≈ [σ2]iP]]$;
    \item[$-$] for any $[[Γ ⊢ iN]]$, $[[Γ' ⊢ [σ1]iN ≈ [σ2]iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by induction on $[[iP]]$ (and mutually on $[[N]]$).
  \begin{caseof}
    \item $[[iN = α⁻]]$\\
      Then since by inversion, $[[α⁻ ∊ {Γ}]]$, 
      $[[Γ' ⊢ [σ1]α⁻ ≈ [σ2]α⁻]]$ holds by definition of $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$.
    \item $[[iN = ↑iP]]$\\
      Then by inversion, $[[Γ ⊢ iP]]$.
      By the induction hypothesis, $[[Γ' ⊢ [σ1]iP ≈ [σ2]iP]]$, 
      Then by \ruleref{\ottdruleDOneShiftULabel}, 
      $[[Γ' ⊢ ↑[σ1]iP ≤ ↑[σ2]iP]]$, and symmetrically, $[[Γ' ⊢ ↑[σ2]iP ≤ ↑[σ1]iP]]$,
      together meaning that $[[Γ' ⊢ ↑[σ1]iP ≈ ↑[σ2]iP]]$,
      or equivalently, $[[Γ' ⊢ [σ1]↑iP ≈ [σ2]↑iP]]$.
    \item $[[iN = iP → iM]]$\\
      Then by inversion, $[[Γ ⊢ iP]]$ and $[[Γ ⊢ iM]]$.
      By the induction hypothesis, $[[Γ' ⊢ [σ1]iP ≈ [σ2]iP]]$ and
      $[[Γ' ⊢ [σ1]iM ≈ [σ2]iM]]$,
      that is $[[Γ' ⊢ [σ1]iP ≥ [σ2]iP]]$, $[[Γ' ⊢ [σ2]iP ≥ [σ1]iP]]$,
      $[[Γ' ⊢ [σ1]iM ≤ [σ2]iM]]$, and $[[Γ' ⊢ [σ2]iM ≤ [σ1]iM]]$.
      Then by \ruleref{\ottdruleDOneArrowLabel}, $[[Γ' ⊢ [σ1]iP → [σ1]iM ≤ [σ2]iP → [σ2]iM]]$, 
      and again by \ruleref{\ottdruleDOneArrowLabel}, $[[Γ' ⊢ [σ2]iP → [σ2]iM ≤ [σ1]iP → [σ1]iM]]$.
      This way, $[[Γ' ⊢ [σ1]iP → [σ1]iM ≈ [σ2]iP → [σ2]iM]]$,
      or equivalently, $[[Γ' ⊢ [σ1](iP → iM) ≈ [σ2](iP → iM)]]$.
    \item $[[iN = ∀pas.iM]]$
      We can assume that $[[pas]]$ is disjoint from $[[Γ]]$ and $[[Γ']]$.
      By inversion, $[[Γ ⊢ ∀pas.iM]]$ implies $[[Γ, pas ⊢ iM]]$.
      Notice that $[[Γ' ⊢ σi : Γ]]$ and 
      $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$
      can be extended to 
      $[[Γ', pas ⊢ σi : Γ, pas]]$
      and 
      $[[Γ', pas ⊢ σ1 ≈ σ2 : Γ, pas]]$
      by \cref{lemma:subst-domain-weakening}.
      Then by the induction hypothesis,
      $[[Γ', pas ⊢ [σ1]iM ≈ [σ2]iM]]$,
      meaning by inversion that
      $[[Γ', pas ⊢ [σ1]iM ≤ [σ2]iM]]$ and 
      $[[Γ', pas ⊢ [σ2]iM ≤ [σ1]iM]]$.

@@

\begin{corollary}[Algorithmic substitution equivalence becomes equality after normalization]
  \label{corollary:subst-subt-equiv-algorithmization}
  Suppose that $[[Θ ⊢ uσ1]]$ and $[[Θ ⊢ uσ2]]$ are algorithmic substitutions.
  Then
  $[[Θ ⊢ uσ1 ≈ uσ2]] \iff [[nf(uσ1)|dom(Θ) = nf(uσ2)|dom(Θ)]]$
  and 
  $[[Γ ⊢ uσ1 ≈ uσ2 : Ξ]] \iff [[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$.
\end{corollary}

>>>>

\begin{corollary}[Normalization preserves the signature of the algorithmic substitution]
  \label{corollary:norm-subst-sig-algo}
  \hfill \\
  $[[Θ ⊢ uσ : Ξ]] \iff [[Θ ⊢ nf(uσ) : Ξ]]$,
  $[[Γ ⊢ uσ : Ξ]] \iff [[Γ ⊢ nf(uσ) : Ξ]]$.
\end{corollary}
\begin{proof}
  The proof is analogous to \cref{corollary:wf-s-nf}.
\end{proof}

@@

\begin{lemma}[Equivalent substitution act equivalently]
  \label{lemma:equiv-subst-on-same-term}
  Suppose that $[[Γ' ⊢ σ1 : Γ]]$ and $[[Γ' ⊢ σ2 : Γ]]$
  are substitutions equivalent on their domain, that is $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$.
  Then
  \begin{itemize}
    \item[$+$] for any $[[Γ ⊢ iP]]$, $[[Γ' ⊢ [σ1]iP ≈ [σ2]iP]]$;
    \item[$-$] for any $[[Γ ⊢ iN]]$, $[[Γ' ⊢ [σ1]iN ≈ [σ2]iN]]$.
  \end{itemize}
\end{lemma}
\begin{proof}
  We prove it by induction on $[[iP]]$ (and mutually on $[[iN]]$).
  \begin{caseof}
    \item $[[iN = α⁻]]$\\
      Then since by inversion, $[[α⁻ ∊ {Γ}]]$, 
      $[[Γ' ⊢ [σ1]α⁻ ≈ [σ2]α⁻]]$ holds by definition of $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$.
    \item $[[iN = ↑iP]]$\\
      Then by inversion, $[[Γ ⊢ iP]]$.
      By the induction hypothesis, $[[Γ' ⊢ [σ1]iP ≈ [σ2]iP]]$, 
      Then by \ruleref{\ottdruleDOneShiftULabel}, 
      $[[Γ' ⊢ ↑[σ1]iP ≤ ↑[σ2]iP]]$, and symmetrically, $[[Γ' ⊢ ↑[σ2]iP ≤ ↑[σ1]iP]]$,
      together meaning that $[[Γ' ⊢ ↑[σ1]iP ≈ ↑[σ2]iP]]$,
      or equivalently, $[[Γ' ⊢ [σ1]↑iP ≈ [σ2]↑iP]]$.
    \item $[[iN = iP → iM]]$\\
      Then by inversion, $[[Γ ⊢ iP]]$ and $[[Γ ⊢ iM]]$.
      By the induction hypothesis, $[[Γ' ⊢ [σ1]iP ≈ [σ2]iP]]$ and
      $[[Γ' ⊢ [σ1]iM ≈ [σ2]iM]]$,
      that is $[[Γ' ⊢ [σ1]iP ≥ [σ2]iP]]$, $[[Γ' ⊢ [σ2]iP ≥ [σ1]iP]]$,
      $[[Γ' ⊢ [σ1]iM ≤ [σ2]iM]]$, and $[[Γ' ⊢ [σ2]iM ≤ [σ1]iM]]$.
      Then by \ruleref{\ottdruleDOneArrowLabel}, $[[Γ' ⊢ [σ1]iP → [σ1]iM ≤ [σ2]iP → [σ2]iM]]$, 
      and again by \ruleref{\ottdruleDOneArrowLabel}, $[[Γ' ⊢ [σ2]iP → [σ2]iM ≤ [σ1]iP → [σ1]iM]]$.
      This way, $[[Γ' ⊢ [σ1]iP → [σ1]iM ≈ [σ2]iP → [σ2]iM]]$,
      or equivalently, $[[Γ' ⊢ [σ1](iP → iM) ≈ [σ2](iP → iM)]]$.
    \item $[[iN = ∀pas.iM]]$
      We can assume that $[[pas]]$ is disjoint from $[[Γ]]$ and $[[Γ']]$.
      By inversion, $[[Γ ⊢ ∀pas.iM]]$ implies $[[Γ, pas ⊢ iM]]$.
      Notice that $[[Γ' ⊢ σi : Γ]]$ and 
      $[[Γ' ⊢ σ1 ≈ σ2 : Γ]]$
      can be extended to 
      $[[Γ', pas ⊢ σi : Γ, pas]]$
      and 
      $[[Γ', pas ⊢ σ1 ≈ σ2 : Γ, pas]]$
      by \cref{lemma:subst-domain-weakening}.
      Then by the induction hypothesis,
      $[[Γ', pas ⊢ [σ1]iM ≈ [σ2]iM]]$,
      meaning by inversion that
      $[[Γ', pas ⊢ [σ1]iM ≤ [σ2]iM]]$ and 
      $[[Γ', pas ⊢ [σ2]iM ≤ [σ1]iM]]$.

@@

\begin{corollary}[Algorithmic substitution equivalence becomes equality after normalization]
  \label{corollary:subst-subt-equiv-algorithmization}
  Suppose that $[[Θ ⊢ uσ1 : Ξ']]$ and $[[Θ ⊢ uσ2 : Ξ']]$ are algorithmic substitutions
  and $[[Ξ ⊆ Ξ']]$. 
  Then $[[Γ ⊢ uσ1 ≈ uσ2 : Ξ]] \iff [[nf(uσ1)|Ξ = nf(uσ2)|Ξ]]$.
\end{corollary}

### 2023-09-01 proofs/tex/unif-constraint-merge-lemmas.tex
Finish syntax update

\begin{corollary}
    \label{corollary:unif-merge-soundness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$ 
    are normalized unification constraints.
    If $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined then
    \begin{enumerate}
        \item $[[Θ ⊢ UC]]$ is normalized unification constraint,
        \item for any substitution $[[Θ ⊢ uσ]]$, $[[ Θ   ⊢ uσ : lift UC ]]$ implies 
        $[[ Θ   ⊢ uσ : lift UC1 ]]$ and $[[ Θ   ⊢ uσ : lift UC2 ]]$.
    \end{enumerate}
\end{corollary}
\begin{proof}
    It is clear that since $[[UC = UC1 ∪ UC2]]$ (by \cref{lemma:unif-merge-soundness}),
    and being normalized means that all entries are normalized,
    $[[UC]]$ is a normalized unification constraint.
    Analogously, $[[Θ ⊢ UC]] = [[UC1 ∪ UC2]]$ holds immediately, 
    since $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.

    Let us take an arbitrary substitution $[[Θ ⊢ uσ]]$ and assume that 
    $[[ Θ   ⊢ uσ : lift UC ]]$.
    Then $[[ Θ   ⊢ uσ : lift UCi ]]$ holds by definition:
    If $[[ucE]] \in [[lift UCi]] \subseteq [[lift UC1 ∪ lift UC2]] = [[lift UC]]$.
    So $[[Θ(α̂±) ⊢ [uσ]α̂± : ucE]]$ holds.
\end{proof}

% \begin{lemma} [Completeness of Unification Constraint Merge]
%     \label{lemma:unif-merge-completeness}
%     Suppose that $[[Θ ⊢ UC|varset1]]$ and $[[Θ ⊢ UC|varset2]]$.
%     Then $[[Θ ⊢ UC|varset1 & UC|varset2 = UC']]$ exists and
%     $[[UC' = UC|varset1 ∪ varset2]]$.
% \end{lemma}
% \begin{proof}
%     $[[Θ ⊢ UC|varset1 & UC|varset2 = UC']]$ is defined as the union of three parts:
%     entries of $[[UC|varset1]]$ that do not have matching entries of $[[UC|varset2]]$,
%     entries of $[[UC|varset2]]$ that do not have matching entries of $[[UC|varset1]]$,
%     and the merge of matching entries.
%     The first two parts are defined. The merge of matching entries is defined
%     by \ruleref{\ottdruleUCMEPEqEqLabel}, since the matching entries must be equal
%     if they both belong to $[[UC]]$.

%     It remains to show that $[[UC' = UC|varset1 ∪ varset2]]$.
%     It is easy to see that the three parts comprising $[[UC']]$ 
%     correspond to the three parts comprising 
%     $[[UC|varset1 ∪ varset2]] = [[UC | (varset1 \ varset2) ∪ UC | (varset2 \ varset1) ∪ UC | varset1 ∩ varset2]]$. 
% \end{proof}

@@

\begin{lemma} [Completeness of Unification Constraint Merge] 
    \label{lemma:unif-merge-completeness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.
    Then for any substitution $[[Θ ⊢ uσ]]$ such that $[[ Θ   ⊢ uσ : lift UC1 ]]$ and $[[ Θ   ⊢ uσ : lift UC2 ]]$, 
    \begin{enumerate}
        \item $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined and
        \item $[[ Θ   ⊢ uσ : UC ]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof repeats the proof of \cref{lemma:merge-completeness} but 
    uses \cref{lemma:unif-entry-merge-completeness} instead of \cref{lemma:entry-merge-completeness}.
\end{proof}

>>>>

\begin{corollary}
    \label{corollary:unif-merge-soundness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$ 
    are normalized unification constraints.
    If $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined then
    \begin{enumerate}
        \item $[[Θ ⊢ UC]]$ is normalized unification constraint,
        \item for any substitution $[[Θ ⊢ uσ : dom(UC)]]$, $[[ Θ  ⊢ uσ : lift UC ]]$ implies 
        $[[ Θ ⊢ uσ : lift UC1 ]]$ and $[[Θ ⊢ uσ : lift UC2]]$.
    \end{enumerate}
\end{corollary}
\begin{proof}
    It is clear that since $[[UC = UC1 ∪ UC2]]$ (by \cref{lemma:unif-merge-soundness}),
    and being normalized means that all entries are normalized,
    $[[UC]]$ is a normalized unification constraint.
    Analogously, $[[Θ ⊢ UC]] = [[UC1 ∪ UC2]]$ holds immediately, 
    since $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.

    Let us take an arbitrary substitution $[[Θ ⊢ uσ : dom(UC)]]$ and assume that 
    $[[ Θ   ⊢ uσ : lift UC ]]$.
    Then $[[ Θ   ⊢ uσ : lift UCi ]]$ holds by definition:
    If $[[ucE]] \in [[lift UCi]] \subseteq [[lift UC1 ∪ lift UC2]] = [[lift UC]]$ then
    $[[Θ(α̂±) ⊢ [uσ]α̂± : ucE]]$ (where $[[ucE]]$ restricts $[[α̂±]]$) holds since $[[Θ ⊢ uσ : dom(UC)]]$.
\end{proof}

@@

\begin{lemma} [Completeness of Unification Constraint Merge] 
    \label{lemma:unif-merge-completeness}
    Suppose that $[[Θ ⊢ UC1]]$ and $[[Θ ⊢ UC2]]$.
    Then for any $[[Ξ]] \supseteq [[dom(UC1) ∪ dom(UC2)]]$ 
    and substitution $[[Θ ⊢ uσ : Ξ]]$ such that $[[ Θ ⊢ uσ : lift UC1 ]]$ and $[[ Θ   ⊢ uσ : lift UC2 ]]$, 
    \begin{enumerate}
        \item $[[Θ ⊢ UC1 & UC2 = UC]]$ is defined and
        \item $[[ Θ   ⊢ uσ : UC ]]$.
    \end{enumerate}
\end{lemma}
\begin{proof}
    The proof repeats the proof of \cref{lemma:merge-completeness} but 
    uses \cref{lemma:unif-entry-merge-completeness} instead of \cref{lemma:entry-merge-completeness}.
\end{proof}
